---
title: "EDA Assignment Prep"
author: "Amber Camp"
format: html
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
library(dplyr)
```

### Data

```{r}
pokedata <- read_csv(here("data/pokemon.csv"))

pokedata <- pokedata %>%
  clean_names()
```

### Describe the data:

Data author: Alberto Barradas

Description: This is pokemon data that has the name, type, and statistics of 800 pokemon looking at 11 different stats.

How many observations (rows) are there? 800

How many variables (columns) are there? 13

link: <https://www.kaggle.com/datasets/abcsds/pokemon>

### skimr: Quick overview

```{r}
skimdata_poke <- skimr:: skim(pokedata)
```

### Summary: Quick overview

```{r}
summary(pokedata)
```

### Categorical vs. Numeric

```{r}
class(pokedata$number)
pokedata$number <- as.factor(pokedata$number)

summary(pokedata$type_1)
pokedata$type_1 <- as.factor(pokedata$type_1)
class(pokedata$type_1)
summary(pokedata$type_1)

summary(pokedata$attack)
```

How many categorical variables are there? 6

How many continuous variables are there? 7

What are some variables of interest to you?

------------------------------------------------------------------------

## EDA Assignment Prep: \[National Park species data\]

### Libraries

### Data

```{r}
np_species <- read_csv("data/most_visited_nps_species_data.csv")
```

### Describe the data:

Data author: by National Park Service uploaded by **Umer Haddii**

Description: Species identification and sightings in US national parks

How many observations (rows) are there? 61119 observations

How many variables (columns) are there? 28 variables

### Summary: Quick overview

```{r}
summary(np_species)

```

### Categorical vs. Numeric

```{r}
class(np_species$Observations)
np_species$Observations <- as.factor(np_species$Observations)

summary(np_species$ParkName)
np_species$ParkName <- as.factor(np_species$ParkName)
class(np_species$ParkName)
summary(np_species$ParkName)

summary(np_species$CategoryName)
```

How many categorical variables are there?

-   15

How many continuous variables are there?

-   13

What are some variables of interest to you?

-   The frequency of which parks had the most wildlife sightings.

What kinds of questions could you explore with this data?

-   What animal was seen the most?

```{r}
specific_parks <- np_species %>%
  filter(ParkName %in% c("Yellowstone National Park",
                         "Grand Canyon National Park",
                         "Joshua Tree National Park"))
```

```{r}
park_amount_of_observations <- ggplot(specific_parks, aes(x = ParkName, y = CategoryName)) +
geom_point()

# Display the basic plot
park_amount_of_observations

# Flip coordinates for horizontal orientation
#park_amount_of_observations + coord_flip()

# Convert to polar coordinates for a circular layout
#park_amount_of_observations + coord_polar()

```

```{r}
filtered_category_data <- specific_parks %>%
  group_by(CategoryName) %>%           # Group by the column of interest
  filter(n() >= 1)      #keep everything over 1

category_counts <- specific_parks %>%
  count(CategoryName, sort = TRUE) 
```

```{r}
ggplot(category_counts, aes(x = CategoryName, y = n)) +
geom_col() +
coord_flip()+ 
  theme_light()
  
```
