---
title: "reading_in_data"
format: html
editor: visual
---

## Read in Data

```{r}
library(tidyverse)
library(plotly)
library(janitor)
library(ggpubr)
```

```{r}
mtcars <- datasets::mtcars
```

```{r}
?mtcars
```

How do we change it so that the row names are the collumn names?

```{r}
mtcars <- mtcars %>%
  rownames_to_column("model")
```

Lets examine the relationship between weight and miles per gallon.

```{r}
cars_wt_mpg <- ggplot(data = mtcars, aes(x = mpg, y = wt, point = model, color = hp)) +
geom_point() +
geom_smooth() +
  stat_cor() +
  labs(title = "weight to Miles per gallon",
    x = "Miles Per Galon",
        y = "Wieght of Vehicle")
cars_wt_mpg
```

```{r}
mtcars_split <- mtcars %>%
separate(model, into = c("make", "model"), sep = " ")
```

you can also assign #label = model does the same as #point = model

```{r}
ggplotly(cars_wt_mpg)

```

#we have to runt he above script before we run our scatter plot because it changed the column name so we know what model car is coresponding to the point.

lets add text to the points we can use #geom_label or #geom_text

```{r}
cars_wt_mpg <- ggplot(data = mtcars, aes(x = mpg, y = wt, point = model, color = hp)) +
geom_point() +
  labs(title = "weight to Miles per gallon",
    x = "Miles Per Galon",
        y = "Wieght of Vehicle")
cars_wt_mpg
```

What is the avergae mpg for each car manufacturer, please plot the results.

```{r}
average_cars_mpg <- mtcars_split %>%
  group_by(make)%>%
  summarise(avg_mpg = mean(mpg))
average_cars_mpg
```

```{r}
mtcars_split %>%
  select(make, mpg) %>%
  summarise(avg_mpg = mean(mpg))
```

```{r}

ggplot(data = average_cars_mpg, aes(x = avg_mpg, y = make, fill = avg_mpg)) +
       geom_bar(stat = "identity")
```
