---
title: "Mapping"
format: html
editor: visual
---

### Overview

Making maps is awesome and there are a ton of options/packages in R we can utiliize

Today, we are going to keep it simple and focus on the following

-   What is an sf object?

-   Gradient vs Discrete Coloring

-   Using familiar packages, `ggplot` and `plotly` to make a world map

### Libraries

```{r}
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
library(scales)
library(plotly)
options(scipen = 99)
```

### Load in World Data

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

unique(world$name_long)
```

### Plotting

#### Static, blank world map

```{r}
ggplot(data = world) +
    geom_sf()
```

#### Gradient filled color map

```{r}
ggplot(data = world, aes(fill = pop_est)) +
    geom_sf()
```

#### Continuous to Discrete

```{r}
world <- world %>%
  mutate(pop_bin = cut(
    pop_est,
    breaks = c(0, 1000000, 10000000, 100000000, 500000000, Inf),  # Define meaningful ranges
    labels = c("<1M", "1M-10M", "10M-100M", "100M-500M", ">500M"),  # Bin labels
    include.lowest = TRUE
  ))
```

```{r}
ggplot(data = world, aes(fill = pop_bin)) +
    geom_sf() +
    scale_fill_viridis_d()
```

#### Interactivity

```{r}
world_pop_map_binned <- ggplot(data = world, aes(fill = pop_bin, label = name_long)) +
    geom_sf() +
    scale_fill_viridis_d() +
  theme_minimal()

ggplotly(world_pop_map_binned)
```

### Challenge

Can you make a binned color map showing gdp by country?

```{r}
{r}
# trying Jenks
# install.packages("classInt")
# library(classInt)

jenks <- classIntervals(world$gdp_md, n = 10, style = "jenks")

jenks$brks

world_jenks <- world %>%
  mutate(jenks = cut(gdp_md, breaks = jenks$brks, include.lowest = TRUE, labels = FALSE))

gdp_world_jenks <- ggplot(data = world_jenks, aes(fill = as.factor(jenks), label = name_long)) +
  geom_sf() +
  scale_fill_viridis_d() +
  labs(fill = "GDP Bins") + 
  theme_minimal()

ggplotly(gdp_world_jenks)

jenks_labels <- c(
  paste0("Bin 1: ", round(jenks$brks[1], 2), " to ", round(jenks$brks[2], 2)),
  paste0("Bin 2: ", round(jenks$brks[2], 2), " to ", round(jenks$brks[3], 2)),
  paste0("Bin 3: ", round(jenks$brks[3], 2), " to ", round(jenks$brks[4], 2)),
  paste0("Bin 4: ", round(jenks$brks[4], 2), " to ", round(jenks$brks[5], 2)),
  paste0("Bin 5: ", round(jenks$brks[5], 2), " to ", round(jenks$brks[6], 2)),
  paste0("Bin 6: ", round(jenks$brks[6], 2), " to ", round(jenks$brks[7], 2)),
  paste0("Bin 7: ", round(jenks$brks[7], 2), " to ", round(jenks$brks[8], 2)),
  paste0("Bin 8: ", round(jenks$brks[8], 2), " to ", round(jenks$brks[9], 2)),
  paste0("Bin 9: ", round(jenks$brks[9], 2), " to ", round(jenks$brks[10], 2)),
  paste0("Bin 10: ", round(jenks$brks[10], 2), " and above")
)

gdp_world_jenks_labeled <- ggplot(data = world_jenks, aes(fill = as.factor(jenks), label=name_long)) +
  geom_sf() +
  scale_fill_viridis_d(labels = jenks_labels) +
  labs(fill = "GDP Bins") +
  theme_minimal()

ggplotly(gdp_world_jenks_labeled)
```
